\section{Common usage}
\label{sec:common_usage}


\subsection{Available COMPSs modules}

COMPSs is configured as a Linux Module. As shown in next Figure, the users can type the \verb|module available COMPSs| command to list 
the supported COMPSs modules in the supercomputer. The users can also execute the \verb|module load COMPSs/<version>| command to load an
specific COMPSs module.

\begin{lstlisting}[language=bash]
$ module available COMPSs
---------- /apps/modules/modulefiles/tools ----------
COMPSs/1.3
COMPSs/1.4
COMPSs/2.0
COMPSs/2.1
COMPSs/2.2

COMPSs/release(default)
COMPSs/trunk


$ module load COMPSs/release
load java/1.8.0u66 (PATH, MANPATH, JAVA_HOME, JAVA_ROOT, JAVA_BINDIR,
                    SDK_HOME, JDK_HOME, JRE_HOME)
load MKL/11.0.1 (LD_LIBRARY_PATH)
load PYTHON/2.7.3 (PATH, MANPATH, LD_LIBRARY_PATH, C_INCLUDE_PATH)
load COMPSs/release (PATH, MANPATH, COMPSS_HOME)
\end{lstlisting}

The following command can be run to check if the correct COMPSs version has been loaded:

\begin{lstlisting}[language=bash]
$ enqueue_compss --version
COMPSs version <version>

\end{lstlisting}


\subsection{Configuration}

The COMPSs module contains \textbf{all} the COMPSs dependencies, including Java, Python and MKL. Modifying any of these dependencies
can cause execution failures and thus, we \textbf{do not} recomend to change them. Before running any COMPSs job please check your 
environment and, if needed, comment out any line inside the \verb|.bashrc| file that loads custom COMPSs, Java, Python and/or MKL
modules.

The COMPSs module needs to be loaded in all the nodes that will run COMPSs jobs. Consequently, the \verb|module load| \textbf{must}
be included in your \verb|.bashrc| file. To do so, please run the following command with the corresponding COMPSs version:

\begin{lstlisting}[language=bash]
$ cat "module load COMPSs/release" >> ~/.bashrc
\end{lstlisting}

Log out and back in again to check that the file has been correctly edited. The next listing shows an example of the
output generated by well loaded COMPSs installation. 

\begin{lstlisting}[language=bash]
$ exit
$ ssh USER@SC
load java/1.8.0u66 (PATH, MANPATH, JAVA_HOME, JAVA_ROOT, JAVA_BINDIR,
                    SDK_HOME, JDK_HOME, JRE_HOME)
load MKL/11.0.1 (LD_LIBRARY_PATH)
load PYTHON/2.7.3 (PATH, MANPATH, LD_LIBRARY_PATH, C_INCLUDE_PATH)
load COMPSs/release (PATH, MANPATH, COMPSS_HOME)

$ enqueue_compss --version
COMPSs version <version>
\end{lstlisting}

\colorComment{Please remember that COMPSs runs in several nodes and your current enviroment is not exported to them. Thus, all the 
needed environment variables \textbf{must} be loaded through the \textit{.bashrc} file.}


\subsection{COMPSs Job submission}

COMPSs jobs can be easily submited by running the \textbf{enqueue\_compss} command. This command allows to configure any 
\textbf{runcompss} option and some particular queue options such as the queue system, the number of nodes, the wallclock time,
the master working directory, the workers working directory and number of tasks per node.

Next, we provide detailed information about the \textit{enqueue\_compss} command:

\begin{lstlisting}[language=bash]
$ enqueue_compss -h

Usage: enqueue_compss [queue_system_options] [COMPSs_options] 
          application_name [application_arguments]

          
* Options:                                                                                                          
  General:                                                                                                          
    --help, -h                              Print this help message                                                 
           
           
  Queue system configuration:                                                                                       
    --sc_cfg=<name>                         SuperComputer configuration file to use. 
                                            Must exist inside queues/cfgs/ 
                                            Default: default                                                        
             
             
  Submission configuration:                                                                                         
    --exec_time=<minutes>                   Expected execution time of the application (in minutes)                 
                                            Default: 10        
                                            
    --num_nodes=<int>                       Number of nodes to use                                                  
                                            Default: 2    
                                            
    --num_switches=<int>                    Maximum number of different switches. 
                                            Select 0 for no restrictions.     
                                            Maximum nodes per switch: 18                                            
                                            Only available for at least 4 nodes.                                    
                                            Default: 0 
                                            
    --queue=<name>                          Queue name to submit the job. Depends on the queue 
                                            system.
                                            For example (Nord3): bsc_cs | bsc_debug | debug 
                                            | interactive
                                            Default: default
                                            
    --reservation=<name>                    Reservation to use when submitting the job. 
                                            Default: disabled
                                            
    --job_dependency=<jobID>                Postpone job execution until the job dependency 
                                            has ended.
                                            Default: None
                                            
    --storage_home=<string>                 Root installation dir of the storage implementation
                                            Default: null
                                            
    --storage_props=<string>                Absolute path of the storage properties file
                                            Mandatory if storage_home is defined
                                            
                                            
  Launch configuration:
    --cpus_per_node=<int>                   Available CPU computing units on each node
                                            Default: 16
                                            
    --gpus_per_node=<int>                   Available GPU computing units on each node
                                            Default: 0
                                            
    --max_tasks_per_node=<int>              Maximum number of simultaneous tasks running on a node
                                            Default: -1
                                            
    --node_memory=<MB>                      Maximum node memory: disabled | <int> (MB)
                                            Default: disabled
                                            
    --network=<name>                        Communication network for transfers: 
                                            default | ethernet | infiniband | data.
                                            Default: infiniband
                                              
    --prolog="<string>"                     Task to execute before launching COMPSs (Notice the
                                            quotes). If the task has arguments split them by ","
                                            rather than spaces. 
                                            This argument can appear multiple times for more 
                                            than one prolog action
                                            Default: Empty
                                            
    --epilog="<string>"                     Task to execute after executing the COMPSs application
                                            (Notice the quotes). If the task has arguments split
                                            them by "," rather than spaces.
                                            This argument can appear multiple times for more 
                                            than one epilog action
                                            Default: Empty

    --master_working_dir=<path>             Working directory of the application
                                            Default: .
                                            
    --worker_working_dir=<name | path>      Worker directory. Use: scratch | gpfs | <path>
                                            Default: scratch
                                              
    --worker_in_master_cpus=<int>           Maximum number of CPU computing units that the 
                                            master node can run as worker. 
                                            Cannot exceed cpus_per_node.
                                            Default: 0
                                            
    --worker_in_master_memory=<int> MB      Maximum memory in master node assigned to the worker. 
                                            Cannot exceed the node_memory.
                                            Mandatory if worker_in_master_tasks is specified.
                                            Default: disabled
                                            
    --jvm_worker_in_master_opts="<string>"  Extra options for the JVM of the COMPSs Worker in 
                                            the Master Node. Each option separed by "," and without
                                            blank spaces (Notice the quotes)
                                            Default: Empty
                                            
    --container_image=<path>                Runs the application by means of a singularity 
                                            container image
                                            Default: Empty
                                            
    --container_compss_path=<path>          Path where compss is installed in the Singularity 
                                            container image
                                            Default: /opt/COMPSs

                                            
  Runcompss configuration:

  Tools enablers:
    --graph=<bool>, --graph, -g             Generation of the complete graph (true/false)
                                            When no value is provided it is set to true
                                            Default: false
                                            
    --tracing=<level>, --tracing, -t        Set generation of traces and/or tracing level 
                                            ( [ true | basic ] | advanced | false)
                                            True and basic levels will produce the same traces.
                                            When no value is provided it is set to true
                                            Default: false
                                            
    --monitoring=<int>, --monitoring, -m    Period between monitoring samples (milliseconds)
                                            When no value is provided it is set to 2000
                                            Default: 0
                                            
    --external_debugger=<int>,
    --external_debugger                     Enables external debugger connection on the 
                                            specified port (or 9999 if empty)
                                            Default: false

  Runtime configuration options:
    --task_execution=<compss|storage>       Task execution under COMPSs or Storage.
                                            Default: compss
                                            
    --storage_conf=<path>                   Path to the storage configuration file
                                            Default: None
                                            
    --project=<path>                        Path to the project XML file
                                            Default: default_project.xml
                                            
    --resources=<path>                      Path to the resources XML file
                                            Default: default_resources.xml
                                            
    --lang=<name>                           Language of the application (java/c/python)
                                            Default: Inferred if possible. Otherwise: java
                                            
    --summary                               Displays a task execution summary at the end of 
                                            the application execution
                                            Default: false
                                            
    --log_level=<level>, --debug, -d        Set the debug level: off | info | debug
                                            Default: off

  Advanced options:
    --extrae_config_file=<path>             Sets a custom extrae config file. Must be in a 
                                            shared disk between all COMPSs workers.
                                            Default: null
                                            
    --comm=<ClassName>                      Class that implements the adaptor for communications
                                            Supported adaptors: 
                                            es.bsc.compss.nio.master.NIOAdaptor 
                                            | es.bsc.compss.gat.master.GATAdaptor
                                            Default: es.bsc.compss.nio.master.NIOAdaptor
                                            
    --conn=<className>                      Class that implements the runtime connector for 
                                            the cloud
                                            Supported connectors: 
                                            es.bsc.compss.connectors.DefaultSSHConnector
                                            Default: es.bsc.compss.connectors.DefaultSSHConnector
                                            
    --scheduler=<className>                 Class that implements the Scheduler for COMPSs
                                            Supported schedulers: 
                                            es.bsc.compss.scheduler.fullGraphScheduler
                                            .FullGraphScheduler
                                            | es.bsc.compss.scheduler.fifoScheduler.FIFOScheduler
                                            | es.bsc.compss.scheduler.resourceEmptyScheduler
                                            .ResourceEmptyScheduler
                                            Default: es.bsc.compss.scheduler.loadBalancingScheduler
                                            .LoadBalancingScheduler
                                            
    --library_path=<path>                   Non-standard directories to search for libraries
                                            (e.g. Java JVM library, Python library, C binding
                                            library)
                                            Default: Working Directory
                                            
    --classpath=<path>                      Path for the application classes / modules
                                            Default: Working Directory
                                            
    --appdir=<path>                         Path for the application class folder.
                                            Default: User home
                                            
    --base_log_dir=<path>                   Base directory to store COMPSs log files 
                                            (a .COMPSs/ folder will be created inside this location)
                                            Default: User home
                                            
    --specific_log_dir=<path>               Use a specific directory to store COMPSs log files (the folder
                                            MUST exist and no sandbox is created)
                                            Warning: Overwrites --base_log_dir option
                                            Default: Disabled
                                            
    --uuid=<int>                            Preset an application UUID
                                            Default: Automatic random generation
                                            
    --master_name=<string>                  Hostname of the node to run the COMPSs master
                                            Default: None
                                            
    --master_port=<int>                     Port to run the COMPSs master communications.
                                            Only for NIO adaptor
                                            Default: [43000,44000]
                                            
    --jvm_master_opts="<string>"            Extra options for the COMPSs Master JVM. Each 
                                            option separed by "," and without blank spaces
                                            (Notice the quotes)
                                            Default: Empty
                                            
    --jvm_workers_opts="<string>"           Extra options for the COMPSs Workers JVMs. Each 
                                            option separed by "," and without blank spaces 
                                            (Notice the quotes)
                                            Default: -Xms1024m,-Xmx1024m,-Xmn400m
                                            
    --cpu_affinity="<string>"               Sets the CPU affinity for the workers
                                            Supported options: disabled, automatic, user defined
                                            map of the form "0-8/9,10,11/12-14,15,16"
                                            Default: automatic
                                            
    --gpu_affinity="<string>"               Sets the GPU affinity for the workers
                                            Supported options: disabled, automatic, user defined
                                            map of the form "0-8/9,10,11/12-14,15,16"
                                            Default: automatic
                                            
    --task_count=<int>                      Only for C/Python Bindings. Maximum number of different
                                            functions/methods, invoked from the application, that 
                                            have been selected as tasks
                                            Default: 50
                                            
    --pythonpath=<path>                     Additional folders or paths to add to the PYTHONPATH
                                            Default: User home
                                            
    --PyObject_serialize=<bool>             Only for Python Binding. Enable the object 
                                            serialization to string when possible (true/false).
                                            Default: false


* Application name:
    For Java applications:   Fully qualified name of the application
    For C applications:      Path to the master binary
    For Python applications: Path to the .py file containing the main program

* Application arguments:
    Command line arguments to pass to the application. Can be empty.                                         
\end{lstlisting}

